What is an Ancestral Recombination Graph?

### 2022-03-02 paper.tex
Improve the caption for fig 4
Use "immediate inheritance" vs "sample resolved annotations".

\begin{figure}
\begin{center}
    \includegraphics[width=\textwidth]{illustrations/arg-in-pedigree}
\end{center}
\caption{\label{fig-arg-in-pedigree}
% TODO caption needs another pass. One the full section has been written
% it'll be clearer what we need to say in the caption vs the text.
The Ancestral Recombination Graph defined in terms of genomes embedded
in a pedigree. (a) Diploid individuals, visualised in a highly inbred pedigree and
labelled $i_1$ to $i_8$, contain both paternal (left) and maternal (right) genomes
labelled \textsf{A} to \textsf{P}. Black lines show transmission paths connecting
genomes in the current generation (\textsf{A} to \textsf{D}) with their ancestors.
Two recombination breakpoints, at positions 2 and 7 mean that genomes \textsf{A}
and \textsf{C} are independent recombinants of the paternal genomes \textsf{E}
and \textsf{F}: the region of genome transmitted through each edge is shown in
these cases. Where genomes merge into a common ancestor, the merged region
is shown darker, until all genomic regions are fully coalesced in genome \textsf{N}
in the oldest generation.
(b) The equivalent genome ARG, where there is a single type of node (representing
a genome) along with annotations on all edges; annotations showing partial
transmission are highlighted in bold. Recombination, such as that causing the
breakpoint at position 7, is associated with two parent genomes (\textsf{E} and 
\textsf{F}) and a child genome (the recombinant, \textsf{C}).
(c) The edge-annotated genome ARG in (b) fully defines the
local relationships along the genome, as shown in the local
or ``marginal'' trees.
}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{illustrations/ancestry-resolution}
\caption{\label{fig-ancestry-resolution}
Implicit vs explicit edge annotations. (a) An gARG equivalent to the encoding
described by Griffiths can be created by setting all edges to span the full
genome, apart from edges between two recombining nodes and their recombinant
child (edge spans in bold). (b) The sample-resolved gARG, where the Hudson
simplification algorithm has shortened edge spans so that they encapsulate only
the upwards transmission of genetic material from the samples A, B, and C;
plotted line widths are proportional to the genomic span of each edge. The
example ARG used here is taken from \citet[][fig. 1]{wiuf1999recombination}.
}
\end{figure}

>>>>

\begin{figure}
\begin{center}
    \includegraphics[width=\textwidth]{illustrations/arg-in-pedigree}
\end{center}
\caption{\label{fig-arg-in-pedigree}
% TODO caption needs another pass. One the full section has been written
% it'll be clearer what we need to say in the caption vs the text.
The Ancestral Recombination Graph defined in terms of genomes embedded
in a pedigree. (a) Diploid individuals, visualised in a highly inbred pedigree and
labelled $i_1$ to $i_8$, contain both paternal (left) and maternal (right) genomes
labelled \textsf{A} to \textsf{P}. Black lines show transmission paths connecting
genomes in the current generation (\textsf{A} to \textsf{D}) with their ancestors.
Two recombination breakpoints, at positions 2 and 7 mean that genomes \textsf{A}
and \textsf{C} are independent recombinants of the paternal genomes \textsf{E}
and \textsf{F}: in these cases the edge annotations have been shown, indicating the
region of genome transmitted. Genomes are shaded such that where, backwards in time,
they merge into a common ancestor, the merged region is darker; this shows that all
genomic regions are fully coalesced in genome \textsf{N} in the oldest generation.
(b) The equivalent genome ARG, where there is a single type of node (representing
a genome) along with annotations on all edges; annotations showing partial
transmission are highlighted in bold. Recombination, such as that causing the
breakpoint at position 7, is associated with two parent genomes (\textsf{E} and 
\textsf{F}) and a child genome (the recombinant, \textsf{C}).
(c) The edge-annotated genome ARG in (b) fully defines the
local relationships along the genome, as shown in the local
or ``marginal'' trees.
}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{illustrations/ancestry-resolution}
\caption{\label{fig-ancestry-resolution}
The \citet[][fig. 1]{wiuf1999recombination} example graph as a gARG, with contrasting
edge annotation schemes.
(a) Edge annotations that reflect immediate inheritance, i.e. transmission of all
genetic material between parent and child. All edges span the full genome from position 0 to 7,
apart from those linking a recombinant and its two parents (bold).
(b) Sample-resolved edge annotations: the Hudson simplification algorithm has been used to
shorten edge spans to reflect only the upwards transmission of genetic
material from the samples \textsf{A}, \textsf{B}, and \textsf{C}. The depiction of the genome within
each node is shaded as in figure \ref{fig-arg-in-pedigree}a to show
coalescing regions of genome back to the grand MRCA genome \textsf{W}.
In this case, using line widths to reflect the genomic span of each edge can
help to visualise the relative importance of different transmission pathways. 
}
\end{figure}
